---
title: "CODA_Example"
author: "JJM"
date: "2024-08-19"
output: html_document
---

#Import#

```{r `message=FALSE`}
library(haven)
library(dplyr)
library(naniar)
library(ggplot2)
library(tidyverse)
library(haven)
library(dplyr)
library(compositions)
library(rms)
#library(robCompositions)
#library(zCompositions)
#library(MASS)
#library(ggtern)
#library(plotly)
#library(ggtern)
#library(Ternary)
#library(car)
library(rgl)

BCS70_main <- read.delim("~/MRC DTP/Analysis_1 Local/HDrive/Main/UKDA-8547-tab/tab/bcs_age46_main.tab")
BCSAvg <- read.delim("~/MRC DTP/Analysis_1 Local/HDrive/PA/tab/bcs_age46_activpal_avg.tab")
# source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\CODA_CompoCentre.R")
source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\CODA_ITS_1.R")
# source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\CODA_BiPlots_Q23_27_01_2022.R")
# source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\CODA_BiPlots_PERCENTDIFF_27_01_2022.R")
# source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\CODA_BiPlots_Q12_27_01_2022.R")
# source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\CODA_BiPlots_MINSDIFF_27_01_2022.R")
source("C:\\Users\\JohnM/Documents\\MRC DTP\\Analysis_1 Local\\R files - Analysis 1\\LR_ILR.R")
source("C:\\Users\\JohnM\\Desktop\\ProPASS\\CODA_Functions\\LR_ILR_6PART.R")
# source("C:\\Users\\JohnM\\Desktop\\ProPASS\\CODA_Functions\\trigrid.R")
```

#Cleaning

```{r}
PA<-BCSAvg[c('BCSID', 'B10AAMVPAH', 'B10AASTDH','B10AATSTEPH','B10AAWWT','B10AASITH')]
PA <- rename(PA, 'mvpa'= 'B10AAMVPAH'
             , 'wear'='B10AAWWT' ,
             'sedent'='B10AASITH', 
             'totalactivity'='B10AATSTEPH',
             'stand'='B10AASTDH',
             'ID'='BCSID', 
)

Covariates<-BCS70_main[c('BCSID', 'B10CMSEX', 'BD10MBMI', 'BD10BMI', 'B10SMOKIG')]

Covariates<- rename(Covariates, ID = BCSID, Sex = B10CMSEX, BMI = BD10MBMI, BMIS = BD10BMI, Smoker = B10SMOKIG)

PA <- PA[!(PA$mvpa<=0),]
PA <- PA[!(PA$wear<=0),]
PA <- PA[!(PA$totalactivity<=0),]
PA <- PA[!(PA$sedent<=0),]
PA <- PA[!(PA$stand<0),]
PA$LIPA <- (PA$totalactivity - PA$mvpa + PA$stand)
PA$SLEEP <- 24 - PA$wear
PA <- PA[!(PA$LIPA<0),]
PA <- PA[!(PA$SLEEP<0),]
PA <- subset(PA, select= -c(stand))


##drop BMI=-8 as insufficient info - line 132
##consider impute BMI<0 from self report

for (z in 1:length(Covariates$BMI)) {
  if (Covariates$BMI[z]==-8) {
    Covariates$BMI[z] <- Covariates$BMIS[z]
  }
}

Covariates <- Covariates[!(Covariates$BMI<0),]

Covariates$BMICats <- 0
for (z in 1:length(Covariates$BMI)) {
  if (Covariates$BMI[z]<25 & !is.na(Covariates$BMI[z])) {
    Covariates$BMICats[z] <- 0
  }
  else if (Covariates$BMI[z]<=29.99 & !is.na(Covariates$BMI[z])) {
    Covariates$BMICats[z] <- 1
  }
  else if (Covariates$BMI[z]>29.99 & !is.na(Covariates$BMI[z]) ) {
    Covariates$BMICats[z] <- 2
  }
}
Covariates$Smoker <- dplyr::recode(Covariates$Smoker, '1'=0, '2'=1, '3'=2, '4'=3)

```

#Merge dataframes and inspect

```{r}
df <- merge(PA, Covariates, by="ID")

df <- subset(df, select=c("ID", "mvpa", "sedent", "LIPA", "SLEEP", "Sex", "BMICats", "Smoker"))

df <- rename(df, 'ID'= 'ID'
             , 'mvpa'='mvpa' ,
             'sb'='sedent', 
             'lipa'='LIPA',
             'sleep'='SLEEP',
             'sex'='Sex', 
             'bmi'='BMICats',
             'smoker'='Smoker'
)
head(df, n=50)
#vis_miss(df)


```
##Visualise Raw data [Raw means will add to some unhelpful spurious volume]
```{r}

dfplot <- df[, 2:5]
#In 1 min increments
for (col_name in colnames(dfplot)) {
  plot <- ggplot(df, aes_string(x = col_name)) +
    geom_histogram(binwidth = 0.06, fill = "blue", color = "black", alpha = 0.7) +
    geom_vline(aes_string(xintercept = paste0("mean(", col_name, ", na.rm = TRUE)")),
               color = "red", linetype = "dashed", size = 1) +
    labs(title = col_name, x = "Hours", y = "Frequency") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
  print(plot)
}


```

##Now Apply aitchison closure to the data (normalise to sum to 1, creating compositional Data) ["Acomping data"]

```{r}
df_compositional <- data.frame(cbind(df$mvpa, df$lipa, df$sb, df$sleep))
  df_compositional <- rename(df_compositional, mvpa=X1
                 , lipa = X2 ,
                 sb=X3, 
                 sleep  = X4
                 )
df_compositional <- (acomp(df_compositional))

head(df_compositional)
```

#The compositional centre of the sample reflects a more appropriate average and sums to 1.

```{r}
CompoCentre <- mean(acomp(df_compositional))
CompoCentre
sum(CompoCentre)
```
##Visualise a 4 part composition in a 3D simplex ## 
##Beyond 4-part compositions requires a Series of ternary diagrams, each with 3 components. This is challenging for reviewers so I just report the compositional centre in the table 1 and use bi-plots below.

```{r}
# Define vertices of a tetrahedron in 3D space
vertices <- rbind(
  c(1, 0, -1/sqrt(2)),   # mvpa
  c(-1, 0, -1/sqrt(2)),  # lipa
  c(0, 1, 1/sqrt(2)),    # sb
  c(0, -1, 1/sqrt(2))    # sleep
)
# Function to map compositional data to the tetrahedron
map_to_tetrahedron <- function(comp) {
  comp[1] * vertices[1, ] +
  comp[2] * vertices[2, ] +
  comp[3] * vertices[3, ] +
  comp[4] * vertices[4, ]
}
# Apply the mapping to all data points
mapped_data <- t(apply(df_compositional, 1, map_to_tetrahedron))
# Plot the tetrahedron
plot3d(vertices, type = "n", xlab = "mvpa", ylab = "lipa", zlab = "sb")
# Add edges of the tetrahedron
segments3d(rbind(vertices[1,], vertices[2,]), col = "black")
segments3d(rbind(vertices[1,], vertices[3,]), col = "black")
segments3d(rbind(vertices[1,], vertices[4,]), col = "black")
segments3d(rbind(vertices[2,], vertices[3,]), col = "black")
segments3d(rbind(vertices[2,], vertices[4,]), col = "black")
segments3d(rbind(vertices[3,], vertices[4,]), col = "black")
# Plot the compositional data points
points3d(mapped_data, col = "red")
# Optionally, add the labels to the vertices
text3d(vertices, texts = c("mvpa", "lipa", "sb", "sleep"), adj = c(1.5, 1.5))

```

##Biplots
#[NB: Can apply a hotelling test/MANOVA pillai trace test to formally test for global differences in n-part composition and j-groups]

```{r} 
############### standard bar plot
df_biplot <- df
df_biplot$bmi <- ifelse(df$bmi == 0, 0, 1)

### compute standard mean 
gms <- by(df_biplot[,2:5],df_biplot$bmi,function(x) colMeans(x))
## bind this into frame
ccgm2 <- do.call(rbind,gms)
par(mfrow=c(1,2), oma=c(0,0.0,0.0,5.0))
barplot(ccgm2[2,],main="Overweight/Obese",col=rainbow(4),axes=TRUE,ylim=c(0,12),border=NA,names.arg="")
barplot(ccgm2[1,],main="Normal",col=rainbow(4),axes=FALSE,ylim=c(-0,12),border=NA,names.arg="")
mtext(expression(bold("BMI categories")),side=1,outer=T,line=-3.25)

####### compute CoDa means
gms <- by(df_biplot[,2:5],df_biplot$bmi,function(x) mean(acomp(x)))
gms <- do.call(rbind,gms)
# Compute compositional difference with respect to overall center [1/4,1/4,1/4,1/4]
#gcenter <- rep(1/4,4)
gcenter<-as.numeric(mean(acomp(df_biplot[,2:5])))
ccgm2 <- t(apply(gms,1,function(x) x/gcenter))
# now centre 
# # Compute compositional difference with respect to overall center [1/4,1/4,1/4,1/4]
#gcenter <- rep(1/4,4)
#ccgm2 <- t(apply(ccgm2,1,function(x) x/gcenter))
ccgm2 <-log(ccgm2)
par(mfrow=c(1,2), oma=c(0,0.0,0.0,5.0))
# barplot(ccgm2[1,],main="Obese",col=rainbow(4),axes=TRUE,ylim=c(-0.4,0.4),border=NA,names.arg="")
# legend(0,-0.17,lty=1,lwd=4,col=rainbow(4),legend=c("SB","LIPA","MVPA","SLEEP"),bty="n",cex=0.85)
# axis(4,at=ccgm2[2,],labels=as.character(round(exp(ccgm2[1,])*sign(ccgm2[1,]),2)),las=2) # added axis with actaul number which give proportion group/mean 
# #mtext("grour behaviour/compositional mean",side=4, outer=T)
# barplot(ccgm2[2,],main="Normal",col=rainbow(4),axes=FALSE,ylim=c(-0.4,0.4),border=NA,names.arg="")
# axis(4,at=ccgm2[2,],labels=as.character(round(exp(ccgm2[2,])*sign(ccgm2[2,]),2)),las=2)


library(RColorBrewer)
# Choose a color palette from RColorBrewer
colors <- brewer.pal(4, "Set2")  # "Set2" is a nice, soft color palette

# Your existing code with the updated color palette
par(mfrow=c(1,2), oma=c(0,0.0,0.0,5.0))

# First barplot for "Obese"
barplot(ccgm2[2,], main="Obese/Overweight", col=colors, axes=TRUE, ylim=c(-0.4,0.4), border=NA, names.arg="")
legend(0, -0.17, lty=1, lwd=4, col=colors, legend=c("MVPA","SB","LIPA","SLEEP"), bty="n", cex=0.85)
axis(4, at=ccgm2[2,], labels=as.character(round(exp(ccgm2[1,]) * sign(ccgm2[2,]), 2)), las=2)

# Second barplot for "Normal"
barplot(ccgm2[1,], main="Normal", col=colors, axes=FALSE, ylim=c(-0.4,0.4), border=NA, names.arg="")
axis(4, at=ccgm2[2,], labels=as.character(round(exp(ccgm2[2,]) * sign(ccgm2[1,]), 2)), las=2)


cat("Normal or Underweight BMI")
df_biplot %>% filter(bmi==0) %>% select(mvpa, sb, lipa, sleep) %>% acomp() %>% mean()
ccgm2[1,]
cat("Overweight/Obese")
df_biplot %>% filter(bmi==1) %>% select(mvpa, sb, lipa, sleep) %>% acomp() %>% mean()
ccgm2[2,]

###############################################################################################################

```

##Isometric Log ratio Transformation

```{r} 
##Place your composition columns in compo
compo <- df[,2:5]

##Created binary outcome - Place your outcome into 'outcome'
df$bmi_binary <- ifelse(df$bmi == 0, 0, 1)

#Place your outcome into 'outcome'
Outcome <-df$bmi_binary

##Bind them
data <- cbind(compo, Outcome)

# First rotation model - # compute ilr 
#order of columns: [1] MVPA [2] SB [3] LIPA [4] SLEEP
head(compo) 

```



```{r}
#Argument 1 - OA data; 
#Argument 2 - Determines which column will be the first coordinate (e.g. MVPA/ALL OTHER COMPONENTS) Zi (model with base as col 1)
##MODEL 1##
Zi<-LR_ILR(compo,1)

#Fit a model with these coordinates the model
model_data<-cbind(Zi,Outcome)
fiti <-glm(Outcome~.,data=model_data, family= binomial)
summary(fiti)


##Transformed to OR##
fiti_OR <- exp(coef(fiti))
confint_OR <- exp(confint(fiti))

cat("
    
    The interpretation is challenging: This rotation has bases(coefficients) 1: MVPA/All other components; 2: SB/Sleep + LIPA;  3: LIPA/Sleep   
    
    -So ' mvpa relative to other components was associated with a lower odds of being in the overweight/obese category in unadjusted models'  
    
    However. Do not interpret the magnitude from this, just the direction and p-value because the units of your mvpa/all others is in ?Log-ratio units? We will infer magnitude from the ITS 
    
    Can't we just backtransform? Yes, in principal it exists (invILR) function. However, what would this mean? The utility of CODA is to understand the importance of 1 component in the totality of the day, respecting the closed nature of the data. A single OR for MVPA would be misleading without ITS and suggest that 'for every hour I take from other things', my risk drops by X which we know is not correct. 
    \n\n
    
    ")

fiti_p_value <- summary(fiti)$coefficients[,4]
Rotation_1_Output <- cbind(fiti_OR, confint_OR, fiti_p_value)
Rotation_1_Output


```
###The subsequent models rotate the base coordinate from MVPA:ALL others   to  SB:All others; LIPA:All others ; Sleep:All others 
##This is outlined in the below chunks but could be automated with a function.

```{r}
#Argument 1 -PA data;
#Argument 2 -Determines which column will be the first coordinate (e.g. in this chunk, SB/ALL OTHER COMPONENTS) Zii (model with base as col 2, SB)
##MODEL 2##
Zii<-LR_ILR(compo,2)

#Fit a model with these coordinates the model
model_data<-cbind(Zii,Outcome)
fitii <-glm(Outcome~.,data=model_data, family= binomial)
summary(fitii)


##Transformed to OR##
fitii_OR <- exp(coef(fitii))
confint_OR <- exp(confint(fitii))

cat("
    
    The first coordinate of this model looks directly at SB/All others
    
    \n\n
    
    ")

fitii_p_value <- summary(fitii)$coefficients[,4]
Rotation_2_Output <- cbind(fitii_OR, confint_OR, fitii_p_value)
Rotation_2_Output
```

```{r}
#Repeated for the final 2 components##
##MODEL 3##
Ziii<-LR_ILR(compo,3) #LIPA/All Others#

#Fit a model with these coordinates the model
model_data<-cbind(Ziii,Outcome)
fitiii <-glm(Outcome~.,data=model_data, family= binomial)
summary(fitiii)


##Transformed to OR##
fitiii_OR <- exp(coef(fitiii))
confint_OR <- exp(confint(fitiii))

cat("
    
    The first coordinate of this model looks directly at LIPA/All others
    
    \n\n
    
    ")

fitiii_p_value <- summary(fitiii)$coefficients[,4]
Rotation_3_Output <- cbind(fitiii_OR, confint_OR, fitiii_p_value)
Rotation_3_Output



##MODEL 4##
Ziv<-LR_ILR(compo,3) #LIPA/All Others#

#Fit a model with these coordinates the model
model_data<-cbind(Ziv,Outcome)
fitiv <-glm(Outcome~.,data=model_data, family= binomial)
summary(fitiv)


##Transformed to OR##
fitiv_OR <- exp(coef(fitiv))
confint_OR <- exp(confint(fitiv))

cat("
    
    The first coordinate of this model looks directly at SLEEP/All others
    
    \n\n
    
    ")

fitiv_p_value <- summary(fitiv)$coefficients[,4]
Rotation_4_Output <- cbind(fitiv_OR, confint_OR, fitiv_p_value)
Rotation_4_Output

```

##Well done - we've completed table 2.. Just repeat your models with any covariates.

##Isotemporal substitution (MVPA in place of SB)

```{r}

##Here's an Isotemporal Substitution function, feel free to amend as required#
##Return compo to your raw numbers, non-compositional
compo <- df[, 2:5]
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_mvpa_sb <- ITS(df, fiti, 1, 2)
predict_mvpa_lipa <-  ITS(df, fiti, 1, 3)
predict_mvpa_sleep <-  ITS(df, fiti, 1, 4)


##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_mvpa <- mean(acomp(compo))[1]*60*24
closest_index <- which.min(abs(predict_mvpa_sb$mvpa - mean_mvpa))
ITS_centre <- predict_mvpa_sb[closest_index, ]

# Create a plot using ggplot2
ggplot(predict_mvpa_sb, aes(x = mvpa, y = Prediction)) +
  geom_line(color = "blue") +  # Line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Reference line for time difference = 0
  geom_point(data = ITS_centre, aes(x = mvpa, y = Prediction), color = "black", size = 3) +  # Dot at composition center
  labs(
    title = "",
    x = "Time replaced (minutes)",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  xlim(range(predict_mvpa_sb$mvpa))

```

##Isotemporal Substitution (All behaviours in place of mvpa)

```{r}

compo <- df[, 2:5]
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_mvpa_sb <- ITS(df, fiti, 1, 2)
predict_mvpa_lipa <-  ITS(df, fiti, 1, 3)
predict_mvpa_sleep <-  ITS(df, fiti, 1, 4)


##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_mvpa <- mean(acomp(compo))[1]*60*24
closest_index <- which.min(abs(predict_mvpa_sb$mvpa - mean_mvpa))
ITS_centre <- predict_mvpa_sb[closest_index, ]

# Add a grouping variable to each data frame
predict_mvpa_sb$group <- "Sedentary Behaviour"
predict_mvpa_lipa$group <- "Light intensity PA"
predict_mvpa_sleep$group <- "Sleep"

# Combine all data frames into one
combined_predictions <- rbind(predict_mvpa_sb, predict_mvpa_lipa, predict_mvpa_sleep)

combined_predictions <- combined_predictions %>%
  mutate(
    Prediction = Prediction / ITS_centre$Prediction,
    CI_Lower = CI_Lower / ITS_centre$Prediction,
    CI_Upper = CI_Upper / ITS_centre$Prediction
  )
ITS_centre_rescaled <- ITS_centre
ITS_centre_rescaled$Prediction <- 1

##Choose a sensible x axis limit based (+/- 100 on above sleep average)
x_lims <- c(ITS_centre$mvpa-100, ITS_centre$mvpa+100)

ggplot(combined_predictions, aes(x = mvpa, y = Prediction, color = group)) +
  geom_line() +  # Line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper, fill = group), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Reference line for mvpa = 0
  geom_point(data = ITS_centre_rescaled, aes(x = mvpa, y = Prediction), color = "black", size = 3) +  # Dot at composition center
  labs(
    title = "",
    x = "MVPA (minutes)",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 100)) +  # Fixed x-axis limits from 0 to 100
  scale_color_manual(values = c("blue", "green", "purple")) +  # Set custom colors for each group
  scale_fill_manual(values = c("blue", "green", "purple"))  # Set fill colors matching line colors

```


##Isotemporal substitution (All behaviours in place of sleep)

```{r}
compo <- df[, 2:5]
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_sleep_mvpa <- ITS(df, fiti, 1, 4)
predict_sleep_lipa <-  ITS(df, fiti, 2, 4)
predict_sleep_sb <-  ITS(df, fiti, 3, 4)


##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_sleep <- mean(acomp(compo))[4]*60*24
closest_index <- which.min(abs(predict_sleep_mvpa$sleep - mean_sleep))
ITS_centre <- predict_sleep_mvpa[closest_index, ]


# Add a grouping variable to each data frame
predict_sleep_mvpa$group <- "Moderate-to-vigorous PA"
predict_sleep_lipa$group <- "Light intensity PA"
predict_sleep_sb$group <- "Sedentary Behaviour"

# Combine all data frames into one
combined_predictions <- rbind(predict_sleep_mvpa, predict_sleep_lipa, predict_sleep_sb)

combined_predictions <- combined_predictions %>%
  mutate(
    Prediction = Prediction / ITS_centre$Prediction,
    CI_Lower = CI_Lower / ITS_centre$Prediction,
    CI_Upper = CI_Upper / ITS_centre$Prediction
  )
ITS_centre_rescaled <- ITS_centre
ITS_centre_rescaled$Prediction <- 1

##Choose a sensible x axis limit based (+/- 100 on above sleep average)
x_lims <- c(ITS_centre$sleep-100, ITS_centre$sleep+100)

ggplot(combined_predictions, aes(x = sleep, y = Prediction, color = group)) +
  geom_line() +  # Line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper, fill = group), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Reference line for mvpa = 0
  geom_point(data = ITS_centre_rescaled, aes(x = sleep, y = Prediction), color = "black", size = 3) +  # Dot at composition center
  labs(
    title = "",
    x = "Sleep (minutes)",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = x_lims) +  # Fixed x-axis limits from 0 to 100
  scale_color_manual(values = c("blue", "green", "purple")) +  # Set custom colors for each group
  scale_fill_manual(values = c("blue", "green", "purple"))  # Set fill colors matching line colors



```



##Isotemporal substitution (All behaviours in place of LIPA)

```{r}
compo <- df[, 2:5]
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_mvpa_lipa <- ITS(df, fiti, 1, 3)
predict_sb_lipa <-  ITS(df, fiti, 2, 3)
predict_sleep_lipa <-  ITS(df, fiti, 4, 3)

##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_lipa <- mean(acomp(compo))[3]*60*24
closest_index <- which.min(abs(predict_mvpa_lipa$lipa - mean_lipa))
ITS_centre <- predict_mvpa_lipa[closest_index, ]

# Add a grouping variable to each data frame
predict_mvpa_lipa$group <- "Moderate-to-vigorous PA"
predict_sb_lipa$group <- "Sedentary Behaviour"
predict_sleep_lipa$group <- "Sleep"

# Combine all data frames into one
combined_predictions <- rbind(predict_mvpa_lipa, predict_sb_lipa, predict_sleep_lipa)

combined_predictions <- combined_predictions %>%
  mutate(
    Prediction = Prediction / ITS_centre$Prediction,
    CI_Lower = CI_Lower / ITS_centre$Prediction,
    CI_Upper = CI_Upper / ITS_centre$Prediction
  )
ITS_centre_rescaled <- ITS_centre
ITS_centre_rescaled$Prediction <- 1
##Choose a sensible x axis limit based (+/- 100 on above sleep average)
x_lims <- c(ITS_centre$lipa-100, ITS_centre$lipa+100)

ggplot(combined_predictions, aes(x = lipa, y = Prediction, color = group)) +
  geom_line() +  # Line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper, fill = group), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Reference line for mvpa = 0
  geom_point(data = ITS_centre_rescaled, aes(x = lipa, y = Prediction), color = "black", size = 3) +  # Dot at composition center
  labs(
    title = "",
    x = "Light Intensity PA (minutes)",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = x_lims) +  # Fixed x-axis limits from 0 to 100
  scale_color_manual(values = c("blue", "green", "purple")) +  # Set custom colors for each group
  scale_fill_manual(values = c("blue", "green", "purple"))  # Set fill colors matching line colors

```

##Isotemporal substitution (All behaviours in place of Sedentary behaviour)

```{r}
compo <- df[, 2:5]
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_mvpa_sb <- ITS(df, fiti, 1, 2)
predict_sb_lipa <-  ITS(df, fiti, 2, 3)
predict_sb_sleep <-  ITS(df, fiti, 2, 4)

##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_sb <- mean(acomp(compo))[2]*60*24
closest_index <- which.min(abs(predict_mvpa_sb$sb - mean_sb))
ITS_centre <- predict_mvpa_sb[closest_index, ]
ITS_centre
# Add a grouping variable to each data frame
predict_mvpa_sb$group <- "Moderate-to-vigorous PA"
predict_sb_lipa$group <- "Light intensity PA"
predict_sb_sleep$group <- "Sleep"

# Combine all data frames into one
combined_predictions <- rbind(predict_mvpa_sb, predict_sb_lipa, predict_sb_sleep)

combined_predictions <- combined_predictions %>%
  mutate(
    Prediction = Prediction / ITS_centre$Prediction,
    CI_Lower = CI_Lower / ITS_centre$Prediction,
    CI_Upper = CI_Upper / ITS_centre$Prediction
  )
ITS_centre_rescaled <- ITS_centre
ITS_centre_rescaled$Prediction <- 1
head(combined_predictions)
##Choose a sensible x axis limit based (+/- 100 on above sleep average)
x_lims <- c(ITS_centre$sb-100, ITS_centre$sb+100)

ggplot(combined_predictions, aes(x = sb, y = Prediction, color = group)) +
  geom_line() +  # Line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper, fill = group), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Reference line for mvpa = 0
  geom_point(data = ITS_centre_rescaled, aes(x = sb, y = Prediction), color = "black", size = 3) +  # Dot at composition center
  labs(
    title = "",
    x = "Sedentary Behaviour (minutes)",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = x_lims) +  # Fixed x-axis limits from 0 to 100
  scale_color_manual(values = c("blue", "green", "purple")) +  # Set custom colors for each group
  scale_fill_manual(values = c("blue", "green", "purple"))  # Set fill colors matching line colors


```


```{r}

compo <- df[, 2:5]
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_mvpa_sb <- ITS(df, fiti, 1, 2)
predict_sb_lipa <-  ITS(df, fiti, 2, 3)
predict_sb_sleep <-  ITS(df, fiti, 2, 4)

##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_sb <- mean(acomp(compo))[2]*60*24
closest_index <- which.min(abs(predict_mvpa_sb$sb - mean_sb))
ITS_centre <- predict_mvpa_sb[closest_index, ]
ITS_centre
# Add a grouping variable to each data frame
predict_mvpa_sb$group <- "Moderate-to-vigorous PA"
predict_sb_lipa$group <- "Light intensity PA"
predict_sb_sleep$group <- "Sleep"

# Combine all data frames into one
combined_predictions <- rbind(predict_mvpa_sb, predict_sb_lipa, predict_sb_sleep)
head(combined_predictions)
##Choose a sensible x axis limit based (+/- 100 on above sleep average)
x_lims <- c(ITS_centre$sb-100, ITS_centre$sb+100)

combined_predictions <- combined_predictions %>%
  mutate(
    Prediction = Prediction / ITS_centre$Prediction,
    CI_Lower = CI_Lower / ITS_centre$Prediction,
    CI_Upper = CI_Upper / ITS_centre$Prediction
  )
ITS_centre_rescaled <- ITS_centre
ITS_centre_rescaled$Prediction <- 1
# Enhanced ggplot with white background and dotted grid lines
ggplot(combined_predictions, aes(x = sb, y = Prediction, color = group)) +
  geom_line(size = 1.2) +  # Thicker line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper, fill = group), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", size = 0.5) +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dotted", color = "black", size = 0.8) +  # Dotted reference line for mvpa = 0
  geom_point(data = ITS_centre_rescaled, aes(x = sb, y = Prediction), color = "black", size = 3, shape = 21, fill = "white") +  # Dot at composition center with white fill
  labs(
    title = "",
    x = "Sedentary Behaviour (minutes)",
    y = "Odds Ratio"
  ) +
  scale_x_continuous(limits = x_lims) +  # Fixed x-axis limits
  scale_y_continuous(breaks = seq(0.8, 1.4, by = 0.1)) +  # Set y-axis to increments of 0.1
  scale_color_manual(values = c("blue", "green", "purple")) +  # Custom colors for each group
  scale_fill_manual(values = c("blue", "green", "purple")) +  # Fill colors matching line colors
  theme_classic() +  # White background theme
  theme(
    panel.grid.major.y = element_line(color = "grey80", linetype = "dotted"),  # Dotted grid lines for major y-axis
    panel.grid.minor.y = element_blank(),  # No minor y-axis grid lines
    panel.grid.major.x = element_blank(),  # No x-axis grid lines
    panel.grid.minor.x = element_blank(),  # No x-axis grid lines
    axis.text = element_text(size = 12),  # Adjust axis text size
    axis.title = element_text(size = 14),  # Adjust axis title size
    legend.position = "top"  # Position legend at the top
  )


```






###CODA 5-PART GLM ##---------------------------------------------

```{r}
#Create fictional VPA and MPA#
n <- 5004  # Replace with your desired number of rows
# Generate random mvpa values (e.g., between 30 and 150 minutes)
set.seed(123)  # Set a seed for reproducibility

df1 <- data.frame(mvpa = runif(n, min = 30, max = 150))
# Generate a random proportion for vpa and mpa for each row
random_proportion <- runif(n, min = 0.2, max = 0.5)  # Adjust range as needed
# Calculate vpa and mpa based on random proportions
df$vpa <- df1$mvpa * random_proportion
df$mpa <- df1$mvpa - df$vpa


##Place your composition columns in compo
compo <- df %>% select(mpa, vpa, sleep, sb, lipa)

##Created binary outcome - Place your outcome into 'outcome'
df$bmi_binary <- ifelse(df$bmi == 0, 0, 1)

#Place your outcome into 'outcome'
Outcome <-df$bmi_binary

##Bind them
data <- cbind(compo, Outcome)

# First rotation model - # compute ilr 
#order of columns: [1] MPA [2] VPA [3] SB [4] LIPA [5] SLEEP
head(compo) 

```



```{r}
#Argument 1 - OA data; 
#Argument 2 - Determines which column will be the first coordinate (e.g. MVPA/ALL OTHER COMPONENTS) Zi (model with base as col 1)
##MODEL 1##
Zi<-LR_ILR_5part(compo,1)

#Fit a model with these coordinates the model##
model_data<-cbind(Zi,Outcome)
fiti <-glm(Outcome~.,data=model_data, family= binomial)
summary(fiti)

```

##Example ITS 5-part

```{r}
head(compo)
column_names <- colnames(compo)
predict_sb_vpa <- ITS_5_part(df, fiti, 3, 2)
predict_sb_vpa
```
###Plot 5-part - VPA (col 2) replacing all

```{r}
##ITS requires arguments(df, first_fitteD_model (base_rot=1), column1 in compo to exchange, column2 in compo to exchange between)
predict_vpa_mpa <- ITS_5_part(df, fiti, 2, 1)
predict_vpa_sleep <-  ITS_5_part(df, fiti, 2, 3)
predict_vpa_sb <-  ITS_5_part(df, fiti, 2, 4)
predict_vpa_lipa <-  ITS_5_part(df, fiti, 2, 5)

##Plot the ITS predictions

##Drop a point in at prediction for time exchange ~= 0
mean_vpa <- mean(acomp(compo))[2]*60*24
closest_index <- which.min(abs(predict_vpa_mpa$vpa - mean_vpa))
ITS_centre <- predict_vpa_mpa[closest_index, ]
ITS_centre
# Add a grouping variable to each data frame
predict_vpa_mpa$group <- "mpa"
predict_vpa_sleep$group <- "Sleep"
predict_vpa_sb$group <- "SB"
predict_vpa_lipa$group <- "Lpa"

# Combine all data frames into one
combined_predictions <- rbind(predict_vpa_sleep, predict_vpa_sb, predict_vpa_lipa, predict_vpa_mpa)

combined_predictions <- combined_predictions %>%
  mutate(
    Prediction = Prediction / ITS_centre$Prediction,
    CI_Lower = CI_Lower / ITS_centre$Prediction,
    CI_Upper = CI_Upper / ITS_centre$Prediction
  )

ITS_centre_rescaled <- ITS_centre
ITS_centre_rescaled$Prediction <- 1
head(combined_predictions)
##Choose a sensible x axis limit based (+/- 100 on above sleep average)
x_lims <- c(ITS_centre$vpa-100, ITS_centre$vpa+100)



ggplot(combined_predictions, aes(x = vpa, y = Prediction, color = group)) +
  geom_line() +  # Line for predictions
  geom_ribbon(aes(ymin = CI_Lower, ymax = CI_Upper, fill = group), alpha = 0.2) +  # Shaded area for 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line for OR = 1
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +  # Reference line for mvpa = 0
  geom_point(data = ITS_centre_rescaled, aes(x = vpa, y = Prediction), color = "black", size = 3) +  # Dot at composition center
  labs(
    title = "",
    x = "VPA (minutes)",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  scale_x_continuous(limits = x_lims) +  # Fixed x-axis limits from 0 to 100
  scale_color_manual(values = c("blue", "green", "purple", "red")) +  # Set custom colors for each group
  scale_fill_manual(values = c("blue", "green", "purple", "red"))  # Set fill colors matching line colors


```